---
title: "p8105_hw1_gl2761"
author: "Gonghao Liu"
date: "9/22/2022"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
```

## Problem 1

Load the penguin data set

```{r}
data("penguins", package = "palmerpenguins")
```

#### Data description:
The names of variables include: `r names(penguins)`.  
The important values of variables: `r summary(penguins)`.  
The penguins dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns.  
The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`.  

Making a scatterplot and save it as a .png file.

```{r}

# Make a scatter plot
penguins_plot = ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

# Save the scatter plot to project directory
ggsave("penguins scatter plot.png", penguins_plot)

```

## Problem 2

Creating a data frame.

```{r}

# Set a seed to make sure the randomly generated values are same and reproducible

set.seed(6)

# Create a data frame

problem2_df = tibble(
  var_num = rnorm(10),
  var_logical = if_else(var_num > 0, TRUE, FALSE),
  var_char = c("One", "Two", "Three", "Four", "Five", "Six", "Seven", "Eight", "Nine", "Ten"),
  var_factor = factor(c("A", "B", "C", "A", "B", "C", "A", "B", "C", "A"))
)

# Call data frame

problem2_df

```

Extracting each variable and calculate means

```{r}

# Extract each variable and calculate means

mean(pull(problem2_df, var_num))
# The mean of the variable "var_num" is 0.1053582

mean(pull(problem2_df, var_logical))
# The mean of the variable "var_logical" is 0.7

mean(pull(problem2_df, var_char))
# 'NA' value returned

mean(pull(problem2_df, var_factor))
# 'NA' value returned

```

Using as.numeric function to convert logical, character and factor variables, while only logical and factor variables worked.

```{r, eval = FALSE}

# Convert variable to numerical variables and calculate the mean

# Converting character variable to numeric variable
convert_char = as.numeric(pull(problem2_df, var_char))
# Call convert_char
convert_char
# All elements are "NA"

# Converting factor variable to numeric variable
convert_factor = as.numeric(pull(problem2_df, var_factor))
# Call convert_factor
convert_factor
# All elements in ext_factor are converted to numeric number "1","2" ,and "3"
# Getting mean
mean(convert_factor)
# Mean = 1.9

# Converting logical variable to numeric variable
convert_logical = as.numeric(pull(problem2_df, var_logical))
# Call convert_logcial
convert_factor
# Converted to "1 and 0", 1 for TRUE and 0 for FALSE
# Getting mean
mean(convert_logical)
# Mean = 0.7

```

These can explain 'mean' function can only be used on numeric variables.